---
title: '03.biocrust_IntactBare_community_comp.Rmd'
project: "Jech et al. Biocrust Cultivation"
author: "Sierra Jech"
lab: "Barger Lab"
date: "2024-10-24"
output: html_document
---

# Purpose
-  Create stacked bar plots at the genus and species level for CP and MJ reference communities.

# Setup
Follows after running code in '02.phyloseq_ASV_Table.Rmd' and obtaining the processed phyloseq object. In addition, the '02.phyloseq_ASV_Table.Rmd' code produced a list of 265 cyanobacteria ASVs as a fasta file which was assigned correct names via Cydrasil and iTol. The new list is called 'data_output/cesucyanoassignments265.csv' and should be merged with cyano ASV dataframes when correct genera names need to be used. 

## Libraries

```{r}
library(phyloseq) # For all bioinformatics analyses post Dada2
#packageVersion('phyloseq')
library(ggplot2)
library(tidyverse)
library(vegan)
library(car) # For Levene Test and anova
library(FSA) # dunnTest for post-hoc
```

## Load Data
```{r}
cesu_16S_ps <- readRDS("data_output/mbcyano_processed_filt_ps.rds")
cyanoassignments <- read.csv("data/cesucyanoassignments265.csv")
cyanoassignments$GS <- paste(cyanoassignments$G_checked, cyanoassignments$S_checked)
qPCR <- read.csv("data/cesuqPCR.csv")
```

# Get to know the data, create labels, delete extra samples and blanks
```{r}
cesu_16S_ps@sam_data@names #colnames for the sample_data information in this phyloseq object
cesu_16S_ps@sam_data[["names"]] # view all samples

# Create a combo label where sample types are maintained but the replicate is not included (project, farm or greenhouse, desert, water, substrate)
cesu_16S_ps@sam_data$comboLabel <- paste(cesu_16S_ps@sam_data$project ,cesu_16S_ps@sam_data$farmGreenhouse, cesu_16S_ps@sam_data$desert, cesu_16S_ps@sam_data$water, cesu_16S_ps@sam_data$shade, cesu_16S_ps@sam_data$priorShade,cesu_16S_ps@sam_data$substrate, sep = ".")
unique(cesu_16S_ps@sam_data$comboLabel)
table(cesu_16S_ps@sam_data$comboLabel)

#Setup. Delete P and WP substrates, delete water regime B
# This means I can compare inocula growth in the greenhouse on no substrate, with and without shade (different shade types I think). I will only have timepoint 2 for mayberry farm. I have two timepoints for the greenhouse. I have control values at the greenhouse and farm at the end of cultivation. 

# Subsetting by sample name (remove all sods & watering regime B & blanks):
# now actually remove them in the ps object
cesu_16S_ps_subset <- subset_samples(cesu_16S_ps, !water %in% c("B"))
cesu_16S_ps_subset <- subset_samples(cesu_16S_ps_subset, !substrate %in% c("P", "WP"))
unique(cesu_16S_ps_subset@sam_data$comboLabel)
table(cesu_16S_ps_subset@sam_data$comboLabel)
# one of the blanks has an incorrect value for substrate and shade so it is separating out from the rest in the combolabels - fix it when you melt
# Contruct other useful labels?
# remove blanks for now
cesu_16S_ps_subset <- subset_samples(cesu_16S_ps_subset, !project %in% c("Blank"))

# check it
cesu_16S_ps_subset@sam_data[["comboLabel"]] # ok

# the initials are still in there
```

# Plot relative and absolute abundance of phylum-level groups across the entire project. Keep all samples without shade for now, but maybe remove later
```{r}
 # factor the label levels for plotting
cesu_16S_ps_subset@sam_data$comboLabel <- factor(cesu_16S_ps_subset@sam_data$comboLabel, levels = c("CP.NA.initial.NA.none.NA.soil", "CP.NA.initial.NA.none.none.soil", "MJ.NA.initial.NA.none.NA.soil", "GH0.NA.CP.3.none.NA.greenhouseSoil", "GH0.NA.CP.3.S.NA.greenhouseSoil", "GH0.NA.MJ.3.none.NA.greenhouseSoil", "GH0.NA.MJ.3.S.NA.greenhouseSoil", "GH1.NA.control.3.none.NA.greenhouseSoil", "GH1.NA.control.3.S.NA.greenhouseSoil", "GH1.NA.CP.3.none.NA.greenhouseSoil", "GH1.NA.CP.3.S.NA.greenhouseSoil", "GH1.NA.MJ.3.none.NA.greenhouseSoil", "GH1.NA.MJ.3.S.NA.greenhouseSoil", "MB2.NA.C.A.none.NA.N", "MB2.NA.C.A.S.NA.JC", "MB2.NA.CP.A.none.NA.N", "MB2.NA.CP.A.S.NA.JC", "MB2.NA.MJ.A.none.NA.N", "MB2.NA.MJ.A.S.NA.JC", "R0.g.CP.NA.NA.none.soil", "R0.g.CP.NA.NA.S.soil", "R0.g.MJ.NA.NA.none.soil", "R0.f.MJ.NA.NA.none.soil", "R0.g.MJ.NA.NA.S.soil", "R0.f.CP.NA.NA.none.soil", "R0.f.CP.NA.NA.S.soil", "R0.f.MJ.NA.NA.S.soil", "R0.control.control.NA.NA.none.soil", "R2.g.CP.NA.NA.none.soil", "R2.f.CP.NA.NA.none.soil", "R2.g.CP.NA.NA.S.soil", "R2.f.CP.NA.NA.S.soil", "R2.g.MJ.NA.NA.none.soil", "R2.f.MJ.NA.NA.none.soil", "R2.g.MJ.NA.NA.S.soil", "R2.f.MJ.NA.NA.S.soil", "R2.control.control.NA.NA.none.soil"))

cesu_16S_ps_subset@sam_data$desert <- factor(cesu_16S_ps_subset@sam_data$desert, levels = c("control", "C", "CP", "MJ"))
#Get count of reads by phyla
table(phyloseq::tax_table(cesu_16S_ps_subset)[, "Phylum"]) #40 Phyla included
# get the total read count per sample
reads_per_sample <- as.data.frame (sample_sums(cesu_16S_ps_subset)) # they are all different because we did not rarefy in this dataset
#copy the rownames to a column
reads_per_sample <- tibble::rownames_to_column(reads_per_sample, "Sample")
# change the name of the second column
colnames(reads_per_sample) <- c("Sample", "totalReads")
nrow(reads_per_sample) # 143 samples are included

# first melt the data out of phyloseq format and into long dataframe
cesu_psmelt <- psmelt(cesu_16S_ps_subset) #this takes some time, patience!

# Get a list of the most abundant Phyla by comboLabel
phyla_summary <- cesu_psmelt %>%
  group_by(comboLabel, Phylum) %>%
  summarise(totalAbund = sum(Abundance)) %>%
  arrange(-totalAbund)

# Get a list of the most abundant Phyla
temp <- cesu_psmelt %>%
  group_by(Phylum) %>%
  summarise(totalAbund = sum(Abundance)) %>%
  arrange(-totalAbund)

temp <- temp[1:9,]

top9Phyla <- temp$Phylum # list that holds the Top9Phyla names
top9Phyla

#subset the psmelt dataframe for only the Top9Phyla
phylaSubset <- cesu_psmelt %>%
  filter(Phylum %in% top9Phyla)

# check that we got the right ones
unique(phylaSubset$Phylum) # yes

# calculate the reads per Phylum, grouped by Sample
phylaSubset_abund <- phylaSubset %>%
  group_by(Sample, Phylum) %>%
  summarise(abund = sum(Abundance))

#make a new column were I divide each of the values by the correct one in the reads_per_sample list
# first merge the reads_per_sample data onto the dataframe
phylaSubset_abund2 <- inner_join(phylaSubset_abund, reads_per_sample, by = "Sample")
# calculate reltive abundance: divide the abund column by the total reads column
phylaSubset_abund2$relAbund <- phylaSubset_abund2$abund / phylaSubset_abund2$totalReads
# check that it does not add to 1 because there should be some in the Other category
total_relAbund <- phylaSubset_abund2 %>%
  group_by(Sample) %>%
  summarise(sum = sum(relAbund))
# add a new column which contains the relative abundance value for the Other category
total_relAbund$other <- (1 - total_relAbund$sum)
# delete the sum column
total_relAbund <- total_relAbund %>%
  dplyr::select(Sample, other)
#add column which is "Other" repeated
total_relAbund$Phylum <- "Other"
#change column header "other" to relAbund
colnames(total_relAbund)[which(names(total_relAbund) == "other")] <- "relAbund"
# select columns to keep in the dataframe we want
phylaSubset_abund2 <- phylaSubset_abund2 %>%
  dplyr::select(Sample, relAbund, Phylum)
# rbind the other values to the phylaSubset_abund2 dataframe
phylaSubset_abund3 <- rbind(phylaSubset_abund2, total_relAbund)

# Now check that they sum to 1
total_relAbund2 <- phylaSubset_abund3 %>%
  group_by(Sample) %>%
  summarise(sum = sum(relAbund)) 
head(total_relAbund2) # YES!!!!

# plot it!
# Plot Relative Abundance
temp2 <- unique(cesu_psmelt %>% dplyr::select(Sample, comboLabel, desert, project, shade))
phylaSubset_abund3 <- left_join(phylaSubset_abund3, temp2, by = "Sample")

# Plot Relative Abundance - basic first glimpse
# phylaSubset_abund3 %>%
#   ggplot(aes(y = relAbund, x = Sample, fill = Phylum)) +
#   geom_bar(stat = "identity", colour = "black", linewidth = 0.25) +
#   labs(x = "", y = "Relative Abundance", fill = "Phyla") +
#   theme_classic()+
#   coord_cartesian(ylim = c(0,1), expand = FALSE)+
#   theme(axis.title = element_text(face = "bold", size = 16),
#         axis.text.x = element_text(size = 12, angle = 90),
#         axis.text.y = element_text(size = 12),
#         legend.text = element_text(face = "italic"),
#         axis.ticks.x = element_blank())

# All together but average (not every replicate)
# first calculate the averages
phylaSubset_abund3_agregate <- phylaSubset_abund3 %>% 
  group_by(comboLabel, Phylum, desert, shade, project) %>%
  summarise(meanRel_abund = mean(relAbund),
            sdRel_abund = sd(relAbund))

phylaSubset_abund3_agregate$comboLabel <- factor(phylaSubset_abund3_agregate$comboLabel, levels = c("CP.NA.initial.NA.none.NA.soil", "CP.NA.initial.NA.none.none.soil", "MJ.NA.initial.NA.none.NA.soil", "GH0.NA.CP.3.none.NA.greenhouseSoil", "GH0.NA.CP.3.S.NA.greenhouseSoil", "GH0.NA.MJ.3.none.NA.greenhouseSoil", "GH0.NA.MJ.3.S.NA.greenhouseSoil", "GH1.NA.control.3.none.NA.greenhouseSoil", "GH1.NA.control.3.S.NA.greenhouseSoil", "GH1.NA.CP.3.none.NA.greenhouseSoil", "GH1.NA.CP.3.S.NA.greenhouseSoil", "GH1.NA.MJ.3.none.NA.greenhouseSoil", "GH1.NA.MJ.3.S.NA.greenhouseSoil", "MB2.NA.C.A.none.NA.N", "MB2.NA.C.A.S.NA.JC", "MB2.NA.CP.A.none.NA.N", "MB2.NA.CP.A.S.NA.JC", "MB2.NA.MJ.A.none.NA.N", "MB2.NA.MJ.A.S.NA.JC", "R0.g.CP.NA.NA.none.soil", "R0.g.CP.NA.NA.S.soil", "R0.g.MJ.NA.NA.none.soil", "R0.f.MJ.NA.NA.none.soil", "R0.g.MJ.NA.NA.S.soil", "R0.f.CP.NA.NA.none.soil", "R0.f.CP.NA.NA.S.soil", "R0.f.MJ.NA.NA.S.soil", "R0.control.control.NA.NA.none.soil", "R2.g.CP.NA.NA.none.soil", "R2.f.CP.NA.NA.none.soil", "R2.g.CP.NA.NA.S.soil", "R2.f.CP.NA.NA.S.soil", "R2.g.MJ.NA.NA.none.soil", "R2.f.MJ.NA.NA.none.soil", "R2.g.MJ.NA.NA.S.soil", "R2.f.MJ.NA.NA.S.soil", "R2.control.control.NA.NA.none.soil"))

# plot
#new_labels <- c("CPinitial" = "Colorado\nPlateau", "MJinitial" = "Mojave\nDesert","SNinitial" = "Sonoran\nDesert")

phylaSubset_abund3_agregate$Phylum <- factor(phylaSubset_abund3_agregate$Phylum, levels = c("Other","Planctomycetota","Verrucomicrobiota","Crenarchaeota", "Chloroflexi", "Actinobacteriota", "Acidobacteriota","Cyanobacteria", "Proteobacteria","Bacteroidota"))

ten_colors_green_remixed <- c('#658354', '#536A89', '#BE6C9C', '#0F74C8', '#4E467A', '#9591BF','#2A659B','#AB3A62', '#E395C2', '#508EBF')

phylaSubset_abund3_agregate %>%
  ggplot(aes(y = meanRel_abund, x = comboLabel, fill = Phylum)) +
  geom_bar(stat = "identity", alpha = 0.8, width = 0.85) +
  scale_fill_manual(values = ten_colors_green_remixed, breaks=c("Bacteroidota", "Proteobacteria","Cyanobacteria", "Acidobacteriota", "Actinobacteriota", "Chloroflexi", "Crenarchaeota", "Verrucomicrobiota", "Planctomycetota", 'Other'))+
  scale_y_continuous(expand = c(0, 0.01), breaks=seq(0, 1, .2)) +
  theme_classic()+
  coord_flip()+
  #facet_grid(rows = vars(comboLabel), scales = "free", space = "free", switch = "y")+ ##labeller = labeller(comboLabel = new_labels
  labs(x = "", y = "Relative Abundance", fill = "")+
  theme(axis.text.x = element_text(size = 12),
        #axis.text.y = element_blank(),
        #axis.ticks.y = element_blank(),
        axis.title = element_text(size = 12),
        #strip.background = element_rect(fill="white", colour=NA),
        #strip.background = element_blank(),
        #strip.text = element_blank(),
        #strip.text = element_text(face="bold", size=12),
        #strip.text.y.left = element_text(angle = 0),
        legend.title = element_text(face = "bold"),
        legend.position = "bottom",
        legend.text = element_text(size = 10),
        plot.margin = margin(0.5,1,0,0, "cm"))+
  guides(fill = guide_legend(nrow = 3, byrow = TRUE))                           

#I did not save this relative abundance plot
```

# Subset for controls and intact biocrusts
```{r}
CP_soils<- phylaSubset_abund3 %>%
  filter(project %in% c("R0","R2")) %>%
  filter(desert == "control")
CP_soils$IS <- "S"

CP_initials <- phylaSubset_abund3 %>% 
  filter(project == "CP") %>% 
  group_by(Phylum)
CP_initials$IS <- "I"

# merge these two together
CP_IS <- rbind(CP_initials, CP_soils)
```


I need to recalculate phylaSubset_abund3 for Mojave phyla before continuing
```{r}
MJ_initials_summary <- phylaSubset_abund3 %>% 
  filter(project == "MJ") %>% 
  group_by(Phylum) %>% 
  summarise(min = min(relAbund),
            max = max(relAbund),
            mean = mean(relAbund),
            sd = sd(relAbund)) %>%
  arrange(desc(mean))

# can use these ones from above again
# CP_soils_summary <- phylaSubset_abund3 %>%
#   filter(project %in% c("R0","R2")) %>%
#   filter(desert == "control") %>%
#   group_by(Phylum) %>%
#   summarise(min = min(relAbund),
#             max = max(relAbund),
#             mean = mean(relAbund),
#             sd = sd(relAbund)) %>%
#   arrange(desc(mean))

# CP_soils<- phylaSubset_abund3 %>%
#   filter(project %in% c("R0","R2")) %>%
#   filter(desert == "control")
# CP_soils$IS <- "S"

MJ_initials <- phylaSubset_abund3 %>% 
  filter(project == "MJ") %>% 
  group_by(Phylum)
MJ_initials$IS <- "I"

# merge these two together so that I can run a t.test for each phylum
MJ_IS <- rbind(MJ_initials, CP_soils)
```


# Genera Level Absolute Abundance
```{r}
# need to correct the names of cyanobacteria genera/species with Cydrasil
head(cesu_psmelt)
# add in the Cydrasil names
cyanoassignments_GS <- cyanoassignments %>% dplyr::select("ASV","GS")
cesupsmelt <- left_join(cesu_psmelt, cyanoassignments_GS, by = c("OTU" = "ASV")) 

# text issues with the names 
cesupsmelt$GS <- trimws(cesupsmelt$GS, which = c("right"))
unique(cesupsmelt$GS) # 50 unique GS levels
# fix some of the names 
cesupsmelt$GS[cesupsmelt$GS == "Chroococidopsis"] <- "Chroococcidiopsis"
# Deal with the names
cesupsmelt$GS_all <- cesupsmelt$GS
cesupsmelt$GS_all <- ifelse(is.na(cesupsmelt$GS_all), paste(cesupsmelt$Genus), cesupsmelt$GS_all)
unique(cesupsmelt$GS_all) # 698 unique GS_all levels
```

Calculate whether there are differences in abundance for intact biocrust and bare soil at the Genus/Species (GS) level.
```{r}
# merge dataframe with total reads in sample
cesupsmelt_abund <- inner_join(cesupsmelt, reads_per_sample, by = "Sample")
# divide the Abundance column by the total reads column
cesupsmelt_abund$relAbund <- cesupsmelt_abund$Abundance / cesupsmelt_abund$totalReads

# CP samples
CP_GS_initials_summary <- cesupsmelt_abund %>% 
  filter(project == "CP") %>% 
  group_by(Phylum, GS_all) %>% 
  summarise(min = min(relAbund),
            max = max(relAbund),
            mean = mean(relAbund),
            sd = sd(relAbund)) %>%
  arrange(desc(mean))

# background soil samples 
CP_GS_soils_summary <- cesupsmelt_abund %>%
  filter(project %in% c("R0","R2")) %>%
  filter(desert == "control") %>%
  group_by(Phylum, GS_all) %>%
  summarise(min = min(relAbund),
            max = max(relAbund),
            mean = mean(relAbund),
            sd = sd(relAbund)) %>%
  arrange(desc(mean))


CP_GS_soils <- cesupsmelt_abund %>%
  filter(project %in% c("R0","R2")) %>%
  filter(desert == "control")
CP_GS_soils$IS <- "S"

CP_GS_initials <- cesupsmelt_abund %>% 
  filter(project == "CP") %>% 
  group_by(Phylum, GS_all)
CP_GS_initials$IS <- "I"

# merge these two together so that I can run a t.test for each phylum
CP_GS_IS <- rbind(CP_GS_initials, CP_GS_soils)
```

Create a genus level relative abundance plot for the 40 most abundant genera for Appendix A
```{r}
# Get a list of the most abundant genera in intact biocrust
genera_summary_Intact <- CP_GS_IS %>%
  filter(IS == "I") %>%
  group_by(GS_all, Phylum) %>%
  summarise(totalRelAbund = sum(relAbund)) %>%
  arrange(-totalRelAbund)

# keep the top 40, remove NA
genera_summary_Intact <- genera_summary_Intact %>% filter(!GS_all %in% c("not identified", "NA"))
temp_Intact_GS <- genera_summary_Intact[1:40,]

top40GeneraIntact <- temp_Intact_GS$GS_all # list that holds the names of the top genera
top40GeneraIntact

#subset the CP_GS_IS dataframe for only the top40GeneraIntact so that you get values for intact and background soil
generaIntactSubset <- CP_GS_IS %>%
  filter(GS_all %in% top40GeneraIntact)

# check that we got the right ones
unique(generaIntactSubset$GS_all) # yes

# calculate the reads per Genera, grouped by Sample
generaIntactSubset_abund <- generaIntactSubset %>%
  group_by(Sample, GS_all, IS) %>%
  summarise(relAbund = sum(relAbund))

# check that it does not add to 1 because there should be some in the Other category
total_relAbund_IS <- generaIntactSubset_abund %>%
  group_by(Sample) %>%
  summarise(sum = sum(relAbund)) #these top 40 only capture about 60-70% of the reads

# add a new column which contains the relative abundance value for the Other category
total_relAbund_IS$other <- (1 - total_relAbund_IS$sum)
# delete the sum column
total_relAbund_IS <- total_relAbund_IS %>%
  dplyr::select(Sample, other)
#add column which is "Other" repeated
total_relAbund_IS$GS_all <- "Other"
#change column header "other" to relAbund
colnames(total_relAbund_IS)[which(names(total_relAbund_IS) == "other")] <- "relAbund"
# select columns to keep in the dataframe we want
generaIntactSubset_abund2 <- generaIntactSubset_abund %>%
  dplyr::select(Sample, relAbund, GS_all)
# rbind the other values to the phylaSubset_abund2 dataframe
generaIntactSubset_abund3 <- rbind(generaIntactSubset_abund2, total_relAbund_IS)

# Now check that they sum to 1
total_relAbund2_IS <- generaIntactSubset_abund3 %>%
  group_by(Sample) %>%
  summarise(sum = sum(relAbund)) 
head(total_relAbund2_IS) # YES!!!!
```

# Plot Relative Abundance
```{r}
#temp2 <- unique(cesu_psmelt %>% dplyr::select(Sample, comboLabel, desert, project, shade)) # already ran this code above
generaIntactSubset_abund3 <- left_join(generaIntactSubset_abund3, temp2, by = "Sample")

#Plot Relative Abundance - basic first glimpse
generaIntactSubset_abund3 %>%
  ggplot(aes(y = relAbund, x = Sample, fill = GS_all)) +
  geom_bar(stat = "identity", colour = "black", linewidth = 0.25) +
  labs(x = "", y = "Relative Abundance", fill = "Genus or species") +
  theme_classic()+
  theme(axis.title = element_text(face = "bold", size = 16),
        axis.text.x = element_text(size = 12, angle = 90),
        axis.text.y = element_text(size = 12),
        legend.text = element_text(face = "italic"),
        axis.ticks.x = element_blank())

# All together but average (not every replicate)
# first calculate the averages
generaIntactSubset_abund3_aggregate <- generaIntactSubset_abund3 %>% 
  group_by(GS_all,desert) %>%
  summarise(meanrelAbund = mean(relAbund),
            sdrelAbund = sd(relAbund))

# This is a bad work around to get the labels correct 
levels(generaIntactSubset_abund3_aggregate$desert)
generaIntactSubset_abund3_aggregate$desert <- factor(generaIntactSubset_abund3_aggregate$desert, levels = c("NA","Intact Biocrust","control"))
generaIntactSubset_abund3_aggregate$desert[is.na(generaIntactSubset_abund3_aggregate$desert)] <- "Intact Biocrust"

# plot
IS_labels <- c("Intact Biocrust" = "Intact Biocrust", "control" = "Background Soil")

generaIntactSubset_abund3_aggregate$GS_all <- factor(generaIntactSubset_abund3_aggregate$GS_all, levels = c("Other","Crustifilum","JGI0001001-H03","Couchioplanes","Arcticibacter","Edaphobaculum","Haliangium","Arizonema","Thermobaculum","Nannocystis","Puia","Archangium","Massilia","Rufibacter","Rhodocytophaga","Chroococcidiopsis","Aurantisolimonas","Flavobacterium","Fibrella","Candidatus Udaeobacter","Candidatus Nitrososphaera","Rubellimicrobium","Microvirga","Aridibacter","Hymenobacter","Sphingomonas","Bryobacter","Undefined Coleofasciculaceae", "Ellin6055","Ferruginibacter","Adhaeribacter","Flavisolibacter","Rubrobacter","Abditibacterium","Spirosoma","RB41", "Blastocatella","Candidatus Nitrocosmicus","Chthoniobacter","Segetibacter","Microcoleus vaginatus"))

forty_colors <- c("#A6CEE3", "#1F78B4", "#B2DF8A", "#33A02C", "#FB9A99", "#E31A1C","#FDBF6F", "#FF7F00", "#CAB2D6", "#6A3D9A", "#FFFF99", "#B15928","dodgerblue2", "#E31A1C", "green4","purple4", "orange", "black", "gold1","skyblue2", "#FB9A99", "palegreen2","#CAB2D6", "orange3", "gray70", "khaki2", "maroon", "orchid1", "deeppink1", "blue1", "steelblue4","darkturquoise", "green1", "yellow4", "yellow3","darkorange4", "lightblue", "brown", "grey90","grey30")

generaIntactSubset_abund3_aggregate %>%
  ggplot(aes(y = meanrelAbund, x = desert, fill = GS_all)) +
  geom_bar(stat = "identity", alpha = 0.8, width = 0.85) +
  geom_col(color = "black", linewidth = 0.25)+
  scale_fill_manual(values = forty_colors, breaks=c("Microcoleus vaginatus", "Segetibacter","Chthoniobacter","Candidatus Nitrocosmicus","Blastocatella","RB41","Spirosoma","Abditibacterium","Rubrobacter","Flavisolibacter","Adhaeribacter","Ferruginibacter","Ellin6055","Undefined Coleofasciculaceae","Bryobacter","Sphingomonas","Hymenobacter","Aridibacter","Microvirga","Rubellimicrobium","Candidatus Nitrososphaera","Candidatus Udaeobacter","Fibrella","Flavobacterium","Aurantisolimonas","Chroococcidiopsis","Rhodocytophaga","Rufibacter","Massilia","Archangium","Puia","Nannocystis","Thermobaculum","Arizonema","Haliangium","Edaphobaculum","Arcticibacter","Couchioplanes","JGI0001001-H03","Crustifilum", "Other"))+
  scale_y_continuous(expand = c(0, 0.01), breaks=seq(0, 1, .2)) +
  scale_x_discrete(labels = IS_labels)+
  theme_classic()+
  coord_flip()+
  #facet_grid(rows = vars(comboLabel), scales = "free", space = "free", switch = "y")+ ##labeller = labeller(comboLabel = new_labels
  labs(x = "", y = "Relative Abundance", fill = "")+
  theme(axis.text = element_text(size = 12),
        #axis.text.y = element_blank(),
        #axis.ticks.y = element_blank(),
        axis.title = element_text(size = 12),
        #strip.background = element_rect(fill="white", colour=NA),
        #strip.background = element_blank(),
        #strip.text = element_blank(),
        #strip.text = element_text(face="bold", size=12),
        #strip.text.y.left = element_text(angle = 0),
        #legend.title = element_text(face = "bold"),
        legend.position = "bottom",
        legend.text = element_text(size = 10, face = "italic"),
        plot.margin = margin(0.5,1,0,0, "cm"))+
  guides(fill = guide_legend(nrow = 8, byrow = TRUE))                           

#save this relative abundance plot
#ggsave("figures/IS_GSall_relAbund_stackedbar.pdf", width = 11, height = 6)
```

# Mojave Plot
```{r}
MJ_GS_initials_summary <- cesupsmelt_abund %>% 
  filter(project == "MJ") %>% 
  group_by(Phylum, GS_all) %>% 
  summarise(min = min(relAbund),
            max = max(relAbund),
            mean = mean(relAbund),
            sd = sd(relAbund)) %>%
  arrange(desc(mean))

# background soil samples 
# keep this the same, use the dataframe already created called CP_GS_soils_summary

# this is also the same
#CP_GS_soils <- cesupsmelt_abund %>%
#  filter(project %in% c("R0","R2")) %>%
#  filter(desert == "control")
#CP_GS_soils$IS <- "S"

MJ_GS_initials <- cesupsmelt_abund %>% 
  filter(project == "MJ") %>% 
  group_by(Phylum, GS_all)
MJ_GS_initials$IS <- "I"

# merge these two together so that I can run a t.test for each phylum
MJ_GS_IS <- rbind(MJ_GS_initials, CP_GS_soils)
```

# Create a genus level relative abundance plot for the 40 most abundant genera for Appendix A 
```{r}
# Get a list of the most abundant genera in intact biocrust
MJ_genera_summary_Intact <- MJ_GS_IS %>%
  filter(IS == "I") %>%
  group_by(GS_all, Phylum) %>%
  summarise(totalRelAbund = sum(relAbund)) %>%
  arrange(-totalRelAbund)

# keep the top 40, remove NA
MJ_genera_summary_Intact <- MJ_genera_summary_Intact %>% filter(!GS_all %in% c("not identified", "NA"))
temp_Intact_GS <- MJ_genera_summary_Intact[1:39,]

MJ_top40GeneraIntact <- temp_Intact_GS$GS_all # list that holds the names of the top genera
MJ_top40GeneraIntact

#subset the CP_GS_IS dataframe for only the top40GeneraIntact so that you get values for intact and background soil
MJgeneraIntactSubset <- MJ_GS_IS %>%
  filter(GS_all %in% MJ_top40GeneraIntact)

# check that we got the right ones
unique(MJgeneraIntactSubset$GS_all) # yes

# calculate the reads per Genera, grouped by Sample
MJgeneraIntactSubset_abund <- MJgeneraIntactSubset %>%
  group_by(Sample, GS_all, IS) %>%
  summarise(relAbund = sum(relAbund))

# check that it does not add to 1 because there should be some in the Other category
MJtotal_relAbund_IS <- MJgeneraIntactSubset_abund %>%
  group_by(Sample) %>%
  summarise(sum = sum(relAbund)) #these top 40 only capture about 60-70% of the reads

# add a new column which contains the relative abundance value for the Other category
MJtotal_relAbund_IS$other <- (1 - MJtotal_relAbund_IS$sum)
# delete the sum column
MJtotal_relAbund_IS <- MJtotal_relAbund_IS %>%
  dplyr::select(Sample, other)
#add column which is "Other" repeated
MJtotal_relAbund_IS$GS_all <- "Other"
#change column header "other" to relAbund
colnames(MJtotal_relAbund_IS)[which(names(MJtotal_relAbund_IS) == "other")] <- "relAbund"
# select columns to keep in the dataframe we want
MJgeneraIntactSubset_abund2 <- MJgeneraIntactSubset_abund %>%
  dplyr::select(Sample, relAbund, GS_all)
# rbind the other values to the phylaSubset_abund2 dataframe
MJgeneraIntactSubset_abund3 <- rbind(MJgeneraIntactSubset_abund2, MJtotal_relAbund_IS)

# Now check that they sum to 1
MJtotal_relAbund2_IS <- MJgeneraIntactSubset_abund3 %>%
  group_by(Sample) %>%
  summarise(sum = sum(relAbund)) 
head(MJtotal_relAbund2_IS) # YES!!!!

# plot it!
# Plot Relative Abundance
#temp2 <- unique(cesu_psmelt %>% dplyr::select(Sample, comboLabel, desert, project, shade)) # already ran this code above
MJgeneraIntactSubset_abund3 <- left_join(MJgeneraIntactSubset_abund3, temp2, by = "Sample")

#Plot Relative Abundance - basic first glimpse
MJgeneraIntactSubset_abund3 %>%
  ggplot(aes(y = relAbund, x = Sample, fill = GS_all)) +
  geom_bar(stat = "identity", colour = "black", linewidth = 0.25) +
  labs(x = "", y = "Relative Abundance", fill = "Genus or species") +
  theme_classic()+
  theme(axis.title = element_text(face = "bold", size = 16),
        axis.text.x = element_text(size = 12, angle = 90),
        axis.text.y = element_text(size = 12),
        legend.text = element_text(face = "italic"),
        axis.ticks.x = element_blank())

# All together but average (not every replicate)
# first calculate the averages
MJgeneraIntactSubset_abund3_aggregate <- MJgeneraIntactSubset_abund3 %>% 
  group_by(GS_all,desert) %>%
  summarise(meanrelAbund = mean(relAbund),
            sdrelAbund = sd(relAbund))

# This is a bad work around to get the labels correct 
unique(MJgeneraIntactSubset_abund3_aggregate$desert)
MJgeneraIntactSubset_abund3_aggregate$desert <- factor(MJgeneraIntactSubset_abund3_aggregate$desert, levels = c("NA","Intact Biocrust","control"))
MJgeneraIntactSubset_abund3_aggregate$desert[is.na(MJgeneraIntactSubset_abund3_aggregate$desert)] <- "Intact Biocrust"
levels(MJgeneraIntactSubset_abund3_aggregate$desert)

# plot
#IS_labels <- c("Intact Biocrust" = "Intact Biocrust", "control" = "Background Soil")

MJ_top40GeneraIntact # this is the correct order 
rev(MJ_top40GeneraIntact) # this is the correct order in reverse
MJgeneraIntactSubset_abund3_aggregate$GS_all <- factor(MJgeneraIntactSubset_abund3_aggregate$GS_all, levels = c("Other", "Sphingomonas", "Bryobacter", "Candidatus Udaeobacter", "Nibrella", "Tolypothrix", "Leptolyngbya", "Couchioplanes", "Tumebacillus", "Archangium", "Aridibacter", "Pycnacronema", "Abditibacterium", "Rubrobacter", "Potamolinea", "Geodermatophilus", "Rhodocytophaga", "Massilia", "Undefined Oscillatoriales", "Adhaeribacter", "Rubellimicrobium", "Arizonema", "Flavisolibacter", "RB41", "Chthoniobacter", "Chroococcidiopsis", "Microvirga", "Segetibacter", "Microcoleus paludosus", "Ellin6055", "Blastocatella", "Allocoleopsis", "Pycnacronema brasiliensis", "Parifilum", "Candidatus Nitrocosmicus", "Funiculus", "Schizothrix cf-calcicola", "Scytonema", "Undefined Coleofasciculaceae", "Microcoleus vaginatus"))

forty_colors <- c("#A6CEE3", "#1F78B4", "#B2DF8A", "#33A02C", "#FB9A99", "#E31A1C","#FDBF6F", "#FF7F00", "#CAB2D6", "#6A3D9A", "#FFFF99", "#B15928","dodgerblue2", "#E31A1C", "green4","purple4", "orange", "black", "gold1","skyblue2", "#FB9A99", "palegreen2","#CAB2D6", "orange3", "gray70", "khaki2", "maroon", "orchid1", "deeppink1", "blue1", "steelblue4","darkturquoise", "green1", "yellow4", "yellow3","darkorange4", "lightblue", "brown", "grey90","grey30")

MJgeneraIntactSubset_abund3_aggregate %>%
  ggplot(aes(y = meanrelAbund, x = desert, fill = GS_all)) +
  geom_bar(stat = "identity", alpha = 0.8, width = 0.85) +
  geom_col(color = "black", linewidth = 0.25)+
  scale_fill_manual(values = forty_colors, breaks=c("Microcoleus vaginatus", "Undefined Coleofasciculaceae", "Scytonema", "Funiculus", "Schizothrix cf-calcicola", "Candidatus Nitrocosmicus", "Parifilum", "Pycnacronema brasiliensis", "Allocoleopsis", "Blastocatella", "Ellin6055", "Microcoleus paludosus", "Segetibacter", "Microvirga", "Chroococcidiopsis", "Chthoniobacter", "RB41", "Flavisolibacter", "Arizonema", "Rubellimicrobium", "Adhaeribacter",  "Undefined Oscillatoriales", "Massilia", "Rhodocytophaga", "Geodermatophilus", "Potamolinea", "Rubrobacter", "Abditibacterium", "Pycnacronema", "Aridibacter", "Archangium", "Tumebacillus", "Couchioplanes", "Leptolyngbya", "Tolypothrix", "Nibrella", "Candidatus Udaeobacter", "Bryobacter", "Sphingomonas","Other"))+
  scale_y_continuous(expand = c(0, 0.01), breaks=seq(0, 1, .2)) +
  scale_x_discrete(labels = IS_labels)+
  theme_classic()+
  coord_flip()+
  #facet_grid(rows = vars(comboLabel), scales = "free", space = "free", switch = "y")+ ##labeller = labeller(comboLabel = new_labels
  labs(x = "", y = "Relative Abundance", fill = "")+
  theme(axis.text = element_text(size = 12),
        #axis.text.y = element_blank(),
        #axis.ticks.y = element_blank(),
        axis.title = element_text(size = 12),
        #strip.background = element_rect(fill="white", colour=NA),
        #strip.background = element_blank(),
        #strip.text = element_blank(),
        #strip.text = element_text(face="bold", size=12),
        #strip.text.y.left = element_text(angle = 0),
        #legend.title = element_text(face = "bold"),
        legend.position = "bottom",
        legend.text = element_text(size = 10, face = "italic"),
        plot.margin = margin(0.5,1,0,0, "cm"))+
  guides(fill = guide_legend(nrow = 8, byrow = TRUE))                           

#save this relative abundance plot
#ggsave("figures/MJ_IS_GSall_relAbund_stackedbar.pdf", width = 11, height = 6)
```

