---
title: "11.restore_soil_chemistry"
project: "Jech et al. Biocrust Cultivation"
author: "Sierra Jech"
lab: "Barger Lab"
date: "2024-10-24"
output: html_document
---

# Purpose
-  investigate restoration outcomes 5 months after inoculation

# Setup
Follows after running code in '02.phyloseq_ASV_Table.Rmd' and obtaining the processed phyloseq object. In addition, the '02.phyloseq_ASV_Table.Rmd' code produced a list of 265 cyanobacteria ASVs as a fasta file which was assigned correct names via Cydrasil and iTol. The new list is called 'data_output/cesucyanoassignments265.csv' and should be merged with cyano ASV dataframes when correct genera names need to be used. 

## Libraries
```{r}
library(phyloseq) # For all bioinformatics analyses post Dada2
#packageVersion('phyloseq')
library(ggplot2)
library(tidyverse)
library(vegan)
```

## Load Data
```{r}
soilchem <- read.csv("data/cesu_soilchemdata.csv")
```

## Subset Data for Restoration
```{r}
res_soilchem <- soilchem %>% filter(phase == "resurrect")
# remove cultivation water B
res_soilchem <- res_soilchem %>% filter(cultivation_water != "B")
# keep only the ones that coorespond to the Dec and May sampling timepoints
res_soilchem <- res_soilchem %>% filter(eps_timepoint %in% c("0","2"))

new_labels <- c("g"="Greenhouse Cultivation", "f"="Field Cultivation")
```

# Initial Plots 
```{r}
# Chlorophyll a
res_soilchem %>% ggplot()+
  geom_boxplot(mapping = aes (x = resurrect_desert_treatment, y = chla, fill = eps_timepoint))+
  theme_classic()+
  scale_fill_discrete(name = "Restortation Timepoint", labels = c("Inoculation", "20 weeks"))+
  #scale_x_discrete(labels = c("R2.CP.g.none" = "Colorado\n Plateau", "R2.CP.g.S" = "", "R2.MJ.g.none" = "Mojave", "R2.MJ.g.S" = "", "R2.CP.f.none" = "Colorado\n Plateau", "R2.CP.f.S" = "", "R2.MJ.f.none" = "Mojave", "R2.MJ.f.S" = ""))+
  facet_wrap(~cultivation_f_g_c, scales = "free_x", labeller = as_labeller(new_labels))+
  labs(x = "", y = "Chlorophyll a (units)")+
  theme(axis.ticks.x = element_blank(),
        #axis.text.x = element_text(hjust = -0.5),
        axis.text = element_text(size = 12),
        strip.text = element_text(size = 12),
        axis.title = element_text(size = 14))

# L EPS
res_soilchem %>% ggplot()+
  geom_boxplot(mapping = aes (x = resurrect_desert_treatment, y = lb_eps, fill = eps_timepoint))+
  theme_classic()+
  scale_fill_discrete(name = "Restortation Timepoint", labels = c("Inoculation", "20 weeks"))+
  #scale_x_discrete(labels = c("R2.CP.g.none" = "Colorado\n Plateau", "R2.CP.g.S" = "", "R2.MJ.g.none" = "Mojave", "R2.MJ.g.S" = "", "R2.CP.f.none" = "Colorado\n Plateau", "R2.CP.f.S" = "", "R2.MJ.f.none" = "Mojave", "R2.MJ.f.S" = ""))+
  facet_wrap(~cultivation_f_g_c, scales = "free_x", labeller = as_labeller(new_labels))+
  labs(x = "", y = "LB EPS (units)")+
  theme(axis.ticks.x = element_blank(),
        #axis.text.x = element_text(hjust = -0.5),
        axis.text = element_text(size = 12),
        strip.text = element_text(size = 12),
        axis.title = element_text(size = 14))

# T EPS
res_soilchem %>% ggplot()+
  geom_boxplot(mapping = aes (x = resurrect_desert_treatment, y = tb_eps, fill = eps_timepoint))+
  theme_classic()+
  scale_fill_discrete(name = "Restortation Timepoint", labels = c("Inoculation", "20 weeks"))+
  #scale_x_discrete(labels = c("R2.CP.g.none" = "Colorado\n Plateau", "R2.CP.g.S" = "", "R2.MJ.g.none" = "Mojave", "R2.MJ.g.S" = "", "R2.CP.f.none" = "Colorado\n Plateau", "R2.CP.f.S" = "", "R2.MJ.f.none" = "Mojave", "R2.MJ.f.S" = ""))+
  facet_wrap(~cultivation_f_g_c, scales = "free_x", labeller = as_labeller(new_labels))+
  labs(x = "", y = "TB EPS (units)")+
  theme(axis.ticks.x = element_blank(),
        #axis.text.x = element_text(hjust = -0.5),
        axis.text = element_text(size = 12),
        strip.text = element_text(size = 12),
        axis.title = element_text(size = 14))

# G EPS
res_soilchem %>% ggplot()+
  geom_boxplot(mapping = aes (x = resurrect_desert_treatment, y = g_eps, fill = eps_timepoint))+
  theme_classic()+
  scale_fill_discrete(name = "Restortation Timepoint", labels = c("Inoculation", "20 weeks"))+
  #scale_x_discrete(labels = c("R2.CP.g.none" = "Colorado\n Plateau", "R2.CP.g.S" = "", "R2.MJ.g.none" = "Mojave", "R2.MJ.g.S" = "", "R2.CP.f.none" = "Colorado\n Plateau", "R2.CP.f.S" = "", "R2.MJ.f.none" = "Mojave", "R2.MJ.f.S" = ""))+
  facet_wrap(~cultivation_f_g_c, scales = "free_x", labeller = as_labeller(new_labels))+
  labs(x = "", y = "G EPS (units)")+
  theme(axis.ticks.x = element_blank(),
        #axis.text.x = element_text(hjust = -0.5),
        axis.text = element_text(size = 12),
        strip.text = element_text(size = 12),
        axis.title = element_text(size = 14))


# Total EPS
res_soilchem %>% ggplot()+
  geom_boxplot(mapping = aes (x = resurrect_desert_treatment, y = total_eps, fill = eps_timepoint))+
  theme_classic()+
  scale_fill_discrete(name = "Restortation Timepoint", labels = c("Inoculation", "20 weeks"))+
  #scale_x_discrete(labels = c("R2.CP.g.none" = "Colorado\n Plateau", "R2.CP.g.S" = "", "R2.MJ.g.none" = "Mojave", "R2.MJ.g.S" = "", "R2.CP.f.none" = "Colorado\n Plateau", "R2.CP.f.S" = "", "R2.MJ.f.none" = "Mojave", "R2.MJ.f.S" = ""))+
  facet_wrap(~cultivation_f_g_c, scales = "free_x", labeller = as_labeller(new_labels))+
  labs(x = "", y = "Total EPS (units)")+
  theme(axis.ticks.x = element_blank(),
        #axis.text.x = element_text(hjust = -0.5),
        axis.text = element_text(size = 12),
        strip.text = element_text(size = 12),
        axis.title = element_text(size = 14))

# slake
res_soilchem %>% 
  filter(eps_timepoint == "2") %>%
  ggplot()+
  geom_boxplot(mapping = aes (x = resurrect_desert_treatment, y = slake))+
  theme_classic()+
  #scale_fill_discrete(name = "Restortation Timepoint", labels = c("Inoculation", "20 weeks"))+
  #scale_x_discrete(labels = c("R2.CP.g.none" = "Colorado\n Plateau", "R2.CP.g.S" = "", "R2.MJ.g.none" = "Mojave", "R2.MJ.g.S" = "", "R2.CP.f.none" = "Colorado\n Plateau", "R2.CP.f.S" = "", "R2.MJ.f.none" = "Mojave", "R2.MJ.f.S" = ""))+
  facet_wrap(~cultivation_f_g_c, scales = "free_x", labeller = as_labeller(new_labels))+
  labs(x = "", y = "Soil Stability (units)")+
  theme(axis.ticks.x = element_blank(),
        #axis.text.x = element_text(hjust = -0.5),
        axis.text = element_text(size = 12),
        strip.text = element_text(size = 12),
        axis.title = element_text(size = 14))

```

## Dot plots of the difference between inoculation and 20 weeks later, perhaps vertically for each treatment
```{r}
soilchem_summary <- res_soilchem %>% 
  group_by(resurrect_desert_treatment, cultivation_f_g_c, eps_timepoint) %>%
  summarise(meanChla = mean(chla, na.rm = TRUE),
            sdChla = sd(chla, na.rm = TRUE),
            meanTotalEPS = mean(total_eps, na.rm = TRUE),
            sdTotalEPS = sd(total_eps, na.rm = TRUE),
            medianSlake = median(slake, na.rm = TRUE),
            sdSlake = sd(slake, na.rm = TRUE))

soilchem_summary_0 <- soilchem_summary %>% filter(eps_timepoint == 0)
soilchem_summary_2 <- soilchem_summary %>% filter(eps_timepoint == 2)
soilchem_summary_diff <- soilchem_summary_0
colnames(soilchem_summary_diff) <- c("resurrect_desert_treatment", "cultivation_f_g_c", "eps_timepoint", "meanChla_0", "sdChla_0", "meanTotalEPS_0", "sdTotalEPS_0", "medianSlake_0", "sdSlake_0")
colnames(soilchem_summary_2) <- c("resurrect_desert_treatment", "cultivation_f_g_c", "eps_timepoint", "meanChla_2", "sdChla_2", "meanTotalEPS_2", "sdTotalEPS_2", "medianSlake_2", "sdSlake_2")
soilchem_summary_2 <- soilchem_summary_2 %>% dplyr::select(!eps_timepoint)
soilchem_summary_diff <- left_join(soilchem_summary_diff, soilchem_summary_2, by = c("resurrect_desert_treatment", "cultivation_f_g_c"))

# calculate the differences
soilchem_summary_diff$Chla_diff <- soilchem_summary_diff$meanChla_2 - soilchem_summary_diff$meanChla_0
soilchem_summary_diff$Chla_diff_sd <- sqrt(((soilchem_summary_diff$sdChla_2)^2)+((soilchem_summary_diff$sdChla_0)^2))

soilchem_summary_diff$TotalEPS_diff <- soilchem_summary_diff$meanTotalEPS_2 - soilchem_summary_diff$meanTotalEPS_0
soilchem_summary_diff$TotalEPS_diff_sd <- sqrt(((soilchem_summary_diff$sdTotalEPS_2)^2)+((soilchem_summary_diff$sdTotalEPS_0)^2))

soilchem_summary_diff$slake_diff <- soilchem_summary_diff$medianSlake_2 - soilchem_summary_diff$medianSlake_0
soilchem_summary_diff$slake_diff_sd <- sqrt(((soilchem_summary_diff$sdSlake_2)^2)+((soilchem_summary_diff$sdSlake_0)^2))

# Plotting 
soilchem_summary_diff$cultivation_f_g_c <- factor(soilchem_summary_diff$cultivation_f_g_c, levels = c("g", "f"))
#cultivation_labels <- c("Field Cultivation", "Greenhouse Cultivation")
names(new_labels) <- c("g","f")

soilchem_summary_diff$cultivation_names <- c("control, no shade", "control, shade", "Colorado Plateau, no shade", "Colorado Plateau, shade", "Mojave, no shade", "Mojave, shade","Colorado Plateau, no shade", "Colorado Plateau, shade", "Mojave, no shade", "Mojave, shade")

soilchem_summary_diff$cultivation_names <- factor(soilchem_summary_diff$cultivation_names, levels = c("control, shade", "control, no shade", "Mojave, shade", "Mojave, no shade", "Colorado Plateau, shade", "Colorado Plateau, no shade"))


# Chlorophyll a
soilchem_summary_diff %>%  
  ggplot() + 
  geom_vline(xintercept = 0, linetype = "dotted")+
  geom_point(mapping = aes(y = cultivation_names, x = Chla_diff))+
  geom_errorbarh(mapping = aes(y = cultivation_names, xmin = Chla_diff-Chla_diff_sd, xmax = Chla_diff+Chla_diff_sd), alpha = 0.5, height = 0.2)+
  labs(x = "Difference in Chlorophyll a", y = "")+
  facet_grid(rows = vars(cultivation_f_g_c), scales = "free", space = "free",switch = "y", labeller = labeller(cultivation_f_g_c = new_labels)) +
  theme_classic()+
  theme(strip.placement = "outside",
        strip.background = element_blank(),
        strip.text = element_text(size = 12, face = "bold"),
        axis.text = element_text(size = 12))


# Total EPS
soilchem_summary_diff %>%  
  ggplot() + 
  geom_vline(xintercept = 0, linetype = "dotted")+
  geom_point(mapping = aes(y = cultivation_names, x = TotalEPS_diff))+
  geom_errorbarh(mapping = aes(y = cultivation_names, xmin = TotalEPS_diff-TotalEPS_diff_sd, xmax = TotalEPS_diff+TotalEPS_diff_sd), alpha = 0.5, height = 0.2)+
  labs(x = "Difference in Total EPS", y = "")+
  facet_grid(rows = vars(cultivation_f_g_c), scales = "free", space = "free",switch = "y", labeller = labeller(cultivation_f_g_c = new_labels)) +
  theme_classic()+
  theme(strip.placement = "outside",
        strip.background = element_blank(),
        strip.text = element_text(size = 12, face = "bold"),
        axis.text = element_text(size = 12))


# Soil Stability
soilchem_summary_diff %>%  
  ggplot() + 
  geom_vline(xintercept = 0, linetype = "dotted")+
  geom_point(mapping = aes(y = cultivation_names, x = slake_diff))+
  geom_errorbarh(mapping = aes(y = cultivation_names, xmin = slake_diff-slake_diff_sd, xmax = slake_diff+slake_diff_sd), alpha = 0.5, height = 0.2)+
  labs(x = "Difference in Soil Stability (slake score)", y = "")+
  facet_grid(rows = vars(cultivation_f_g_c), scales = "free", space = "free",switch = "y", labeller = labeller(cultivation_f_g_c = new_labels)) +
  theme_classic()+
  theme(strip.placement = "outside",
        strip.background = element_blank(),
        strip.text = element_text(size = 12, face = "bold"),
        axis.text = element_text(size = 12))

```


## Plotting without difference
Get the initials
```{r}
soilchem_initials <- soilchem %>% filter(phase == "initial")

soilchem_initials_summary <- soilchem_initials %>% 
  group_by(desert_source) %>% 
  summarise(meanChla = mean(chla),
            sdChla = sd(chla),
            meanTotalEPS = mean(total_eps),
            sdTotalEPS = sd(total_eps))

# Axis Labels 
soilchem_summary$cultivation_f_g_c <- factor(soilchem_summary$cultivation_f_g_c, levels = c("g", "f"))
#cultivation_labels <- c("Field Cultivation", "Greenhouse Cultivation")
names(new_labels) <- c("g","f")

soilchem_summary$cultivation_names <- NA
soilchem_summary$cultivation_names <- c("Colorado Plateau, no shade", "Colorado Plateau, no shade", "Colorado Plateau, shade", "Colorado Plateau, shade", "control, no shade", "control, no shade", "control, shade", "control, shade", "Mojave, no shade", "Mojave, no shade", "Mojave, shade", "Mojave, shade","Colorado Plateau, no shade", "Colorado Plateau, no shade", "Colorado Plateau, shade", "Colorado Plateau, shade", "Mojave, no shade", "Mojave, no shade", "Mojave, shade", "Mojave, shade")

soilchem_summary$cultivation_names <- factor(soilchem_summary$cultivation_names, levels = c("control, shade", "control, no shade", "Mojave, shade", "Mojave, no shade", "Colorado Plateau, shade", "Colorado Plateau, no shade"))


# Chlorophyll a
soilchem_summary %>%  
  #filter(eps_timepoint == 0) %>%
  filter(eps_timepoint == 2) %>%
  ggplot() + 
  geom_point(mapping = aes(y = cultivation_names, x = meanChla))+ # color = eps_timepoint
  geom_errorbarh(mapping = aes(y = cultivation_names, xmin = meanChla-sdChla, xmax = meanChla+sdChla), alpha = 0.5, height = 0.2)+
  labs(x = "Chlorophyll a (ug/g soil)", y = "")+
  facet_grid(rows = vars(cultivation_f_g_c), scales = "free", space = "free", switch = "y", labeller = labeller(cultivation_f_g_c = new_labels)) +
  theme_classic()+
  theme(strip.placement = "outside",
        strip.background = element_blank(),
        strip.text = element_text(size = 12, face = "bold"),
        axis.text = element_text(size = 12)) +
  xlim(0,8)


# Total EPS
soilchem_summary %>% 
  #filter(eps_timepoint == 0) %>%
  filter(eps_timepoint == 2) %>%
  ggplot() + 
  geom_point(mapping = aes(y = cultivation_names, x = meanTotalEPS))+
  geom_errorbarh(mapping = aes(y = cultivation_names, xmin = meanTotalEPS-sdTotalEPS, xmax = meanTotalEPS+sdTotalEPS), alpha = 0.5, height = 0.2)+
  labs(x = "Total EPS (ug/g soil)", y = "")+
  facet_grid(rows = vars(cultivation_f_g_c), scales = "free", space = "free", switch = "y", labeller = labeller(cultivation_f_g_c = new_labels)) +
  theme_classic()+
  theme(strip.placement = "outside",
        strip.background = element_blank(),
        strip.text = element_text(size = 12, face = "bold"),
        axis.text = element_text(size = 12))+
  xlim(0,1400)

# Soil Stability
soilchem_summary %>%  
  #filter(eps_timepoint == 0) %>%
  filter(eps_timepoint == 2) %>%
  ggplot() + 
  geom_point(mapping = aes(y = cultivation_names, x = medianSlake))+
  geom_errorbarh(mapping = aes(y = cultivation_names, xmin = medianSlake-sdSlake, xmax = medianSlake+sdSlake), alpha = 0.5, height = 0.2)+
  labs(x = "Soil Stability (slake score)", y = "")+
  facet_grid(rows = vars(cultivation_f_g_c), scales = "free", space = "free",switch = "y", labeller = labeller(cultivation_f_g_c = new_labels)) +
  theme_classic()+
  theme(strip.placement = "outside",
        strip.background = element_blank(),
        strip.text = element_text(size = 12, face = "bold"),
        axis.text = element_text(size = 12))
```


## Statistical Analysis for Figure in main text ####
subset so not comparing irrelevant types
```{r}
# CP
res_soilchem_CP <- res_soilchem %>% filter(desert_source == "CP")
unique(res_soilchem_CP$resurrect_desert_treatment)
kruskal.test(chla ~ resurrect_desert_treatment, data = res_soilchem_CP)
# Kruskal-Wallis chi-squared = 10.809, df = 3, p-value = 0.0128
dunnTest(chla ~ resurrect_desert_treatment, data = res_soilchem_CP)
# F_CP_J - F_CP_JC -3.0332834 0.002419083 0.01451450
# MJ
res_soilchem_MJ <- res_soilchem %>% filter(desert_source == "MJ")
unique(res_soilchem_MJ$resurrect_desert_treatment)
kruskal.test(chla ~ resurrect_desert_treatment, data = res_soilchem_MJ)
# Kruskal-Wallis chi-squared = 0.55193, df = 3, p-value = 0.9073

# Total EPS
# CP
#res_soilchem_CP <- res_soilchem %>% filter(desert_source == "CP")
#unique(res_soilchem_CP$resurrect_desert_treatment)
kruskal.test(total_eps ~ resurrect_desert_treatment, data = res_soilchem_CP)
# Kruskal-Wallis chi-squared = 2.4538, df = 3, p-value = 0.4837

# MJ
#res_soilchem_MJ <- res_soilchem %>% filter(desert_source == "MJ")
#unique(res_soilchem_MJ$resurrect_desert_treatment)
kruskal.test(total_eps ~ resurrect_desert_treatment, data = res_soilchem_MJ)
# Kruskal-Wallis chi-squared = 3.2477, df = 3, p-value = 0.355

# Slake
# CP
kruskal.test(slake ~ resurrect_desert_treatment, data = res_soilchem_CP)
# Kruskal-Wallis chi-squared = 12.878, df = 3, p-value = 0.004908
dunnTest(slake ~ resurrect_desert_treatment, data = res_soilchem_CP)
# F_CP_J - F_CP_JC  2.88224237 0.003948559 0.01579424
# F_CP_JC - G_CP -2.92738428 0.003418262 0.01709131
# F_CP_JC - G_CP_S -3.06281000 0.002192692 0.01315615
res_soilchem_CP %>% ggplot()+
  geom_boxplot(mapping = aes(x = resurrect_desert_treatment, y = slake))+
  theme_classic() # note field+shade has lower soil stability than the others 

# MJ
kruskal.test(slake ~ resurrect_desert_treatment, data = res_soilchem_MJ)
# Kruskal-Wallis chi-squared = 12.325, df = 3, p-value = 0.006348
dunnTest(slake ~ resurrect_desert_treatment, data = res_soilchem_MJ)
# F_MJ_J - G_MJ_S -3.2505647 0.00115176 0.006910563
```


## Plotting a better way
```{r}
# First calculate values for the controls (uninoculated at Resurrect)
soilchem_resurrectControls <- soilchem %>%
  filter(desert_source == "Control") %>%
  filter(phase == "resurrect")
# Chla
mean(soilchem_resurrectControls$chla, na.rm = T)
sd(soilchem_resurrectControls$chla, na.rm = T)
# Total EPS
mean(soilchem_resurrectControls$total_eps, na.rm = T)
sd(soilchem_resurrectControls$total_eps, na.rm = T)
# Slake
median(soilchem_resurrectControls$slake, na.rm = T)
sd(soilchem_resurrectControls$slake, na.rm = T)

# Calculate the values for the inoculated plots 
soilchem_summary2 <- res_soilchem %>% 
  group_by(resurrect_desert_treatment, desert_source, cultivation_f_g_c, eps_timepoint) %>%
  summarise(meanChla = mean(chla, na.rm = TRUE),
            sdChla = sd(chla, na.rm = TRUE),
            meanTotalEPS = mean(total_eps, na.rm = TRUE),
            sdTotalEPS = sd(total_eps, na.rm = TRUE),
            medianSlake = median(slake, na.rm = TRUE),
            sdSlake = sd(slake, na.rm = TRUE))
  
soilchem_summary2$cultivation_names <- NA
soilchem_summary2$cultivation_names <- c("field, no shade", "field, no shade", "field, shade", "field, shade", "field control, no shade", "field control, no shade", "field control, shade", "field control, shade", "field, no shade", "field, no shade", "field, shade", "field, shade","greenhouse, no shade", "greenhouse, no shade", "greenhouse, shade", "greenhouse, shade", "greenhouse, no shade", "greenhouse, no shade", "greenhouse, shade", "greenhouse, shade")

soilchem_summary2$cultivation_names <- factor(soilchem_summary2$cultivation_names, levels = c("field control, shade", "field control, no shade", "field, shade", "field, no shade", "greenhouse, shade", "greenhouse, no shade"))

soilchem_summary2$desert_source <- factor(soilchem_summary2$desert_source, levels = c("CP", "MJ","mayberry_control"))

desert_labels <- c("CP"="Colorado\nPlateau", "MJ"="Mojave", "mayberry_control" = "Field\nCultivation\nControls")

chla_lines <- as.data.frame(c("CP", "MJ","mayberry_control"))
chla_lines$x <- c(1.52, 3.8, 0.48)
chla_lines$sd <- c(2.13, 3.48, 0.41)
colnames(chla_lines) <- c("desert_source", "x", "sd")
chla_lines$desert_source <- factor(chla_lines$desert_source, levels = c("CP", "MJ", "mayberry_control"))

# Chlorophyll a
soilchem_summary2 %>%  
  #filter(eps_timepoint == 0) %>%
  filter(eps_timepoint == 2) %>%
  ggplot() + 
  geom_point(mapping = aes(y = cultivation_names, x = meanChla))+ # color = eps_timepoint
  geom_errorbarh(mapping = aes(y = cultivation_names, xmin = meanChla-sdChla, xmax = meanChla+sdChla), alpha = 0.5, height = 0.2)+
  geom_vline(data = chla_lines, mapping = aes(xintercept = x), color = "grey30", linetype = "dotted")+
  labs(x = expression(paste("Chlorophyll a (",mu, "g/g soil)")), y = "")+
  facet_grid(rows = vars(desert_source), scales = "free", space = "free", switch = "y", labeller = labeller(desert_source = desert_labels)) +
  scale_y_discrete(labels=c("field control, shade" = bquote(~Amelioration[Uninoculated]~""),
                            "field control, no shade" = "Field Control",
                            "field, shade" = bquote(~Amelioration[Field]~""),
                            "field, no shade" = "Field",
                            "greenhouse, shade" = bquote(~Amelioration[GH]~""),
                            "greenhouse, no shade" = "GH"))+
  #facet_grid(rows = vars(cultivation_f_g_c), scales = "free", space = "free", switch = "y", labeller = labeller(cultivation_f_g_c = new_labels)) +
  theme_classic()+
  theme(strip.placement = "outside",
        strip.background = element_blank(),
        strip.text.y.left = element_text(angle = 0),
        strip.text = element_text(size = 12, face = "bold"),
        axis.text = element_text(size = 12))
  #xlim(-0.5,4.5)

# ggsave("figures/chla_rest2.pdf", width = 7, height = 6)


# Total EPS
eps_lines <- as.data.frame(c("CP", "MJ","mayberry_control"))
eps_lines$x <- c(832, 1057, 774)
eps_lines$sd <- c(274, 275, 156)
colnames(eps_lines) <- c("desert_source", "x", "sd")
eps_lines$desert_source <- factor(eps_lines$desert_source, levels = c("CP", "MJ", "mayberry_control"))

soilchem_summary2 %>% 
  #filter(eps_timepoint == 0) %>%
  filter(eps_timepoint == 2) %>%
  ggplot() + 
  geom_point(mapping = aes(y = cultivation_names, x = meanTotalEPS))+
  geom_errorbarh(mapping = aes(y = cultivation_names, xmin = meanTotalEPS-sdTotalEPS, xmax = meanTotalEPS+sdTotalEPS), alpha = 0.5, height = 0.2)+
  geom_vline(data = eps_lines, mapping = aes(xintercept = x), color = "grey30", linetype = "dotted")+
  labs(x = "Total EPS (ug/g soil)", y = "")+
  facet_grid(rows = vars(desert_source), scales = "free", space = "free", switch = "y", labeller = labeller(desert_source = desert_labels)) +
  scale_y_discrete(labels=c("field control, shade" = bquote(~Amelioration[Uninoculated]~""),
                            "field control, no shade" = "Field Control",
                            "field, shade" = bquote(~Amelioration[Field]~""),
                            "field, no shade" = "Field",
                            "greenhouse, shade" = bquote(~Amelioration[GH]~""),
                            "greenhouse, no shade" = "GH"))+
  theme_classic()+
  theme(strip.placement = "outside",
        strip.background = element_blank(),
        strip.text.y.left = element_text(angle = 0),
        strip.text = element_text(size = 12, face = "bold"),
        axis.text = element_text(size = 12))+
  xlim(0,1400)

#ggsave("figures/totaleps_rest2.pdf", width = 7, height = 6)


# Slake
# We did not measure slake for the reference biocrusts but we did for the controls 
slake_lines <- as.data.frame(c("CP", "MJ","mayberry_control"))
slake_lines$x <- c(NA, NA, 2)
slake_lines$sd <- c(NA, NA, 1.2)
colnames(slake_lines) <- c("desert_source", "x", "sd")
slake_lines$desert_source <- factor(slake_lines$desert_source, levels = c("CP", "MJ", "mayberry_control"))


soilchem_summary2 %>%  
  #filter(eps_timepoint == 0) %>%
  filter(eps_timepoint == 2) %>%
  ggplot() + 
  geom_point(mapping = aes(y = cultivation_names, x = medianSlake))+
  geom_errorbarh(mapping = aes(y = cultivation_names, xmin = medianSlake-sdSlake, xmax = medianSlake+sdSlake), alpha = 0.5, height = 0.2)+
  geom_vline(data = slake_lines, mapping = aes(xintercept = x), color = "grey30", linetype = "dotted")+
  labs(x = "Soil Stability (slake score)", y = "")+
  facet_grid(rows = vars(desert_source), scales = "free", space = "free",switch = "y", labeller = labeller(desert_source = desert_labels)) +
  scale_y_discrete(labels=c("field control, shade" = bquote(~Amelioration[Uninoculated]~""),
                            "field control, no shade" = "Field Control",
                            "field, shade" = bquote(~Amelioration[Field]~""),
                            "field, no shade" = "Field",
                            "greenhouse, shade" = bquote(~Amelioration[GH]~""),
                            "greenhouse, no shade" = "GH"))+
  theme_classic()+
  theme(strip.placement = "outside",
        strip.background = element_blank(),
        strip.text = element_text(size = 12, face = "bold"),
        axis.text = element_text(size = 12))+
  xlim(0,6)

#ggsave("figures/slake_restoration.pdf", width = 5.5, height = 6)
```


## Now a comparison with controls
if desert_source says "mayberry control" then new column gets a "C" otherwise it is inoculated
```{r}
res_soilchem$control <- "N"
res_soilchem$control[res_soilchem$desert_source == "mayberry_control"] <- "C"
unique(res_soilchem$control)
kruskal.test(chla ~ control, data = res_soilchem)
# Kruskal-Wallis chi-squared = 1.0436, df = 1, p-value = 0.307
#
kruskal.test(total_eps ~ control, data = res_soilchem)
# Kruskal-Wallis chi-squared = 3.3478, df = 1, p-value = 0.06729
#
kruskal.test(slake ~ control, data = res_soilchem)
#Kruskal-Wallis chi-squared = 0.012348, df = 1, p-value = 0.9115
```


## Now a comparison with initials
```{r}
soilchem_initials$resurrect_desert_treatment <- "initial"
res_soilchem <- res_soilchem %>% select(!control)
res_soilchem_i <- rbind(res_soilchem, soilchem_initials)
unique(res_soilchem_i$resurrect_desert_treatment)
kruskal.test(chla ~ resurrect_desert_treatment, data = res_soilchem_i)
# Kruskal-Wallis chi-squared = 20.799, df = 10, p-value = 0.02254
dunnTest(chla ~ resurrect_desert_treatment, data = res_soilchem_i)
# 48  F_CP_J - initial -3.37452239 0.000739439 0.04066914
#
kruskal.test(total_eps ~ resurrect_desert_treatment, data = res_soilchem_i)
# Kruskal-Wallis chi-squared = 14.938, df = 10, p-value = 0.1343
#
#There are no reference biocrust values for slake 
plot(soilchem_resurrectControls$slake^2,soilchem_resurrectControls$chla)
plot(soilchem_resurrectControls$slake^2,soilchem_resurrectControls$total_eps)
```


# Other Analyses
```{r}
# the resurrect controls are here:
temp_controls <- soilchem_resurrectControls %>% filter(eps_timepoint %in% c("0","2")) %>% filter(!is.na(chla))
# the resurrect samples are here: 
temp <- res_soilchem %>% 
  group_by(resurrect_desert_treatment, desert_source, cultivation_f_g_c, eps_timepoint)
# 
colnames(temp_controls) == colnames(temp)
# merge them together
df <- rbind(temp_controls, temp)
# and merge the initials too
colnames(soilchem_initials) == colnames(df)
df <- rbind(df, soilchem_initials)
df$resurrect_desert_treatment[which(df$phase == "initial")] <- "initial"


# now make some labels
# CP only
df_CP <- df %>% filter(!desert_source %in% c("MJ", "mayberry_control"))

kruskal.test(chla ~ resurrect_desert_treatment, data = df_CP)
# Kruskal-Wallis chi-squared = 16.29, df = 5, p-value = 0.006063
dunnTest(chla ~ resurrect_desert_treatment, data = df_CP, method = "holm")
# 12  F_CP_J - initial -3.0239232 0.002495198 0.03742797

kruskal.test(total_eps ~ resurrect_desert_treatment, data = df_CP)
# Kruskal-Wallis chi-squared = 4.1702, df = 5, p-value = 0.5252
#dunnTest(total_eps ~ resurrect_desert_treatment, data = df_CP, method = "holm")

kruskal.test(slake ~ resurrect_desert_treatment, data = df_CP)
# Kruskal-Wallis chi-squared = 12.478, df = 4, p-value = 0.01413
dunnTest(slake ~ resurrect_desert_treatment, data = df_CP, method = "holm")
# F_CP_JC - G_CP_S -2.96223909 0.003054105 0.03054105
# F_CP_JC - G_CP -2.91424994 0.003565443 0.03208899
# F_CP_J - F_CP_JC  2.79427708 0.005201589 0.04161271

# MJ only
df_MJ <- df %>% filter(!desert_source %in% c("CP", "mayberry_control", "SN"))

kruskal.test(chla ~ resurrect_desert_treatment, data = df_MJ)
# Kruskal-Wallis chi-squared = 11.591, df = 5, p-value = 0.04084
dunnTest(chla ~ resurrect_desert_treatment, data = df_MJ, method = "holm")
#  no significant comparisons

kruskal.test(total_eps ~ resurrect_desert_treatment, data = df_MJ)
# Kruskal-Wallis chi-squared = 7.8217, df = 5, p-value = 0.1663
dunnTest(total_eps ~ resurrect_desert_treatment, data = df_MJ, method = "holm")

kruskal.test(slake ~ resurrect_desert_treatment, data = df_MJ)
# Kruskal-Wallis chi-squared = 10.023, df = 4, p-value = 0.04004
dunnTest(slake ~ resurrect_desert_treatment, data = df_MJ, method = "holm")
# F_MJ_J - G_MJ_S -2.90831040 0.003633875 0.03633875
```